Synthesizing Symmetric Lenses

### 2018-07-12 papers/SymmetricLenses/main.tex
make change

One convenient way to interconvert between these varied data formats is to
use a \emph{lens}~\cite{Focal2005}---a bidirectional program that can
transform data represented in a format $S$ to a format $T$, and vice versa,
offering ``round tripping'' guarantees that help ensure data is not lost or
corrupted as it is transformed back and forth between different
representations.

@@

However, there is a price for narrowing the synthesis domain so
dramatically: Many real data formats are almost, but not quite,
bijections. More often two related data formats have some data in
common, but also contain some differences. For instance, two related
tables might share four of five columns, but each has a different
fifth column. Alternately, one ad hoc configuration file might include
some file-specific meta data, such as a date, a time or a reference
number, whereas the other does not. Indeed of the 48 benchmark problems
described in Section \ref{sec:evaluation}, all of the benchmarks
taken from Flash Fill~\cite{flashfill} and many of the benchmarks that
synchronize between two ad hoc file formats have this 
characteristic, 13 benchmarks in total. Consequently, searching
only for strict bijections is often not applicable in practice.

@@

In summary, our contributions are:\bcp{The section references do not proceed
  in order, so this doesn't work very well as an outline of the paper.
  Should we try to reorder the contributions?  Write a separate outline?
  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's nice
  when the contributions mirror the paper, I am not jarred when they don't.}
\begin{itemize}
\item We define {\em simple symmetric lenses}, with the
  the useful property that they do not maintain any hidden state
  (\S\ref{sec:overview}).  
  We present a collection of combinators for defining these lenses and
  describe how to 
  calculate their likelihoods (\S\ref{sec:ssl}). 
\item  We show that simple symmetric lenses
  fall between asymmetric lenses~\cite{Focal2005-long2} and full
  symmetric lenses~\cite{symmetric-lenses} in expressiveness
  (\S\ref{sec:related}).  We further show that the class of simple
  lenses can be characterized semantically as the subset of full
  symmetric lenses for which a simple ``forgetfulness'' property holds.
\item We show how to synthesize simple symmetric lenses (including all
  asymmetric lenses as a special case), using a novel application of stochastic
  regular expressions to guide the search process (\S\ref{sec:synthesis}). 
  To support that effort, we show that the star semi-ring equivalences on
  regular 
  expressions remain valid when applied to stochastic regular
  expressions, preserving the probability distributions as well as the languages
  (\S\ref{sec:sre}).
\item We extend the Boomerang implementation with simple symmetric lenses,
  demonstrating that they integrate smoothly with 
  other lens extensions such as quotient lenses~\cite{quotientlenses} and matching
  lenses~\cite{matchinglenses}.  We evaluate our
  implementation on a set of 48 lens synthesis benchmarks drawn from
  data cleaning, view maintenance, 
  and file synchronization tasks.  We demonstrate that we can
  synthesize simple symmetric lenses for all of the benchmarks in
  under 30 seconds (\S\ref{sec:evaluation}). 
\end{itemize}
We close with a discussion of related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).

@@

\paragraph*{Asymmetric Lenses}
Formally, an {\em asymmetric lens} $ \ell : S \Leftrightarrow V$ is a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : V \longrightarrow S \longrightarrow S$ and $\ell.\create : V \longrightarrow S$ satisfying the following laws \cite{Focal2005-long2}:
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}\\
\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}
\end{align*}
Despite being less expressive than symmetric lenses, simple symmetric lenses are
strictly more expressive than classical asymmetric lenses.

>>>>

One convenient way to interconvert between these varied data formats is to
use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that can
transform data represented in a format $S$ to a format $T$, and vice versa,
providing ``round tripping'' guarantees that help ensure data is not lost or
corrupted as it is transformed back and forth between different
representations.

@@

However, there is a price for narrowing the synthesis domain so dramatically:
many synchronization tasks cannot by synthesized! Many real data formats are
almost, but not quite, bijections. More often two related data formats have some
data in common, but also contain some differences. For instance, two related
tables might share four of five columns, but each has a different fifth column.
Alternately, one ad hoc configuration file might include some file-specific meta
data, such as a date, a time or a reference number, whereas the other does not.
Indeed of the 48 benchmark problems described in Section \ref{sec:evaluation},
all of the benchmarks taken from Flash Fill~\cite{flashfill} and many of the
benchmarks that synchronize between two ad hoc file formats have this
characteristic, 13 benchmarks in total. Consequently, searching only for strict
bijections is often not applicable in practice.

@@

In summary, our contributions are:\bcp{The section references do not proceed
  in order, so this doesn't work very well as an outline of the paper.
  Should we try to reorder the contributions?  Write a separate outline?
  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's nice
  when the contributions mirror the paper, I am not jarred when they don't.}
\begin{itemize}
\item We define {\em simple symmetric lenses}, with the
  the useful property that they do not maintain any hidden state
  (\S\ref{sec:overview}).  
  We present a collection of combinators for defining these lenses and
  describe how to 
  calculate their likelihoods (\S\ref{sec:ssl}). 
\item  We show that simple symmetric lenses
  fall between asymmetric lenses~\cite{Focal2005-long} and full
  symmetric lenses~\cite{symmetric-lenses} in expressiveness
  (\S\ref{sec:related}).  We further show that the class of simple
  lenses can be characterized semantically as the subset of full
  symmetric lenses for which a simple ``forgetfulness'' property holds.
\item We show how to synthesize simple symmetric lenses (including all
  asymmetric lenses as a special case), using a novel application of stochastic
  regular expressions to guide the search process (\S\ref{sec:synthesis}). 
  To support that effort, we show that the star semi-ring equivalences on
  regular 
  expressions remain valid when applied to stochastic regular
  expressions, preserving the probability distributions as well as the languages
  (\S\ref{sec:sre}).
\item We extend the Boomerang implementation with simple symmetric lenses,
  demonstrating that they integrate smoothly with 
  other lens extensions such as quotient lenses~\cite{quotientlenses} and matching
  lenses~\cite{matchinglenses}.  We evaluate our
  implementation on a set of 48 lens synthesis benchmarks drawn from
  data cleaning, view maintenance, 
  and file synchronization tasks.  We demonstrate that we can
  synthesize simple symmetric lenses for all of the benchmarks in
  under 30 seconds (\S\ref{sec:evaluation}). 
\end{itemize}
We close with a discussion of related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).

@@

\paragraph*{Asymmetric Lenses}
Formally, an {\em asymmetric lens} $ \ell : S \Leftrightarrow V$ is a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : V \longrightarrow S \longrightarrow S$ and $\ell.\create : V \longrightarrow S$ satisfying the following laws \cite{Focal2005-long}:
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}\\
\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}
\end{align*}
Despite being less expressive than symmetric lenses, simple symmetric lenses are
strictly more expressive than classical asymmetric lenses.
